---
title: "noaa_buoy"
author: "Cheryl Logan"
date: "4/17/2018"
output: html_document
---

Load libraries. Must first install tidyverse package if not already installed using `install.packages("tidyverse")`

```{r}
library(tidyverse) # after install.packages("tidyverse")
library(reshape2) # after install.packages("tidyverse")
library(ggthemes) # after install.packages("ggthemes")
```

Download 2016 noaa data (Station 46092 - MBM1)
http://www.ndbc.noaa.gov/station_history.php?station=46092

Read in M1 buoy data using tidyr function `read_table2`

```{r}
buoy <-read_table2("/Users/loga8761/Google Drive/LoganLab/Data/StillwaterCove_CTD_data/46092h2016.txt")
```

Read in 2016 Stillwater CTD  data using tidyr function `read_csv`
```{r}
CTD <-read_csv("/Users/loga8761/Google Drive/LoganLab/Data/StillwaterCove_CTD_data/SWC.SAT333.040816.091306.csv")
```

Take a look at the data frames
```{r}
head(buoy) # view column headers
head(CTD) # view column headers
```
## Data Wrangling

1. Let's start with `buoy`...

Combine header rows 1 & 2 in `buoy` by overwriting column names but separate using an underscore `_`
```{r}
## overwrite column names
names(buoy) <- str_c(names(buoy),                  ## current header
                  buoy %>% filter(`#YY` == "#yr"), ## row1 -- don't say buoy[1,] 
                  sep = "_")                       ## separate by `_`

## inspect
names(buoy) ## Looks a lot better
```
Clean up a bit more to get rid of the `#`s and the `/`s.

```{r}
names(buoy) <- str_replace_all(names(buoy), "#", "")  # replace `#` with nothing 
names(buoy) <- str_replace_all(names(buoy), "/", "_") # replace `/` with `_`

## inspect to make sure it worked
names(buoy)
```

Remove redundant 2nd row with units

```{r}
buoy <- buoy %>%
  filter(YY_yr != "#yr") # filter out row where YY_yr is equal to "#yr"

head(buoy)
```


Fix date stamp. Unite year, month, and day into a single date column. Select date, temp and wind.

```{r}
buoy <- buoy %>%
  unite(date, c(YY_yr, MM_mo, DD_dy), sep = "-") %>% # concatenate contents of 3 columns into one column to join the two datasets using a unique ID
  mutate(date = as.Date(date, "%Y-%m-%d")) %>%
  select(date, WTMP_degC,WSPD_m_s)
head(buoy)
```

2. Wrangle `CTD`... Fix date stamp. Unite year, month, and day into a single date column. Select date, temp, oxygen, and pH.
```{r}
CTD <- CTD %>%
  rename(date="Local Date + Time") %>%
  separate(date, c('date', 'time'), sep = ' ') %>%
  select(date,Temp_CTD, O_CTD, pH_Int) %>% # get temp, DO, and pH
  mutate(date = as.Date(date, "%m/%d/%y")) 
  
head(CTD)
```
## Look at Water Temp

Look at distrubution of water temperatures

```{r}
# plot buoy temp data
ggplot(buoy, aes(date, WTMP_degC)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90)) + # rotate x-axis labels 90 degrees
  scale_x_date(limits = as.Date(c('2016-05-15','2016-09-01')))

# plot CTD temp data
ggplot(CTD, aes(date, Temp_CTD)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90)) + # rotate x-axis labels 90 degrees
  scale_x_date(limits = as.Date(c('2016-05-15','2016-09-01')))
```


In `buoy`, we see water temperatures of 999C, so here we begin to see that 99 or 999 are missing value codes. We can replace it with NA. Another issue is that the values are showing up as character strings rather than integers. Also need to fix this. Will do that later


Get mean daily water temp in 2016 for buoy and CTD data
```{r}
# get mean daily values for buoy data (ignore NAs)
buoy_daily <- buoy %>%
  mutate(temp_degC = gsub("99.0", NA, WTMP_degC), wind_m.s = gsub("999.0", NA, WSPD_m_s)) %>% # make temp column numeric
  mutate(temp_degC = as.numeric(temp_degC), wind_m.s = as.numeric(wind_m.s)) %>%
  group_by(date) %>%
  summarize(temp_M1 = mean(temp_degC, na.rm=T), wind_M1 = mean(wind_m.s, na.rm=T)) %>% # make
  ungroup()

# get mean daily values for CTD data (ignore NAs)
CTD_daily <- CTD %>%
  group_by(date) %>%
  summarize(temp_CTD = mean(Temp_CTD, na.rm=T), DO_CTD = mean(O_CTD, na.rm=T), pH_CTD = mean(pH_Int, na.rm=T)) %>% # make
  ungroup()


head(buoy_daily)
head(CTD_daily)
```
Plot mean daily temp, DO and pH.
```{r}
CTD_daily_plot <- gather(CTD_daily,"data_type","value",2:4)  # reshape

ggplot(CTD_daily_plot, aes(date,value)) +
  geom_point() +
  facet_grid(data_type ~., scales = "free") +
  theme_bw() +
  ggtitle("Stillwater CTD and M1 buoy data") +
  scale_x_date(date_breaks = "1 month", limits = as.Date(c('2016-05-15','2016-09-01')))

```

Merge buoy and CTD data by date
```{r}
CTD_buoy <- left_join(buoy_daily, CTD_daily, "date")
```

Reshape to use facet plotting
```{r}
CTD_buoy_plot <- CTD_buoy %>%
  gather("data_type","value",2:6)  # reshape

head(CTD_buoy_plot)
```
Plot mean daily wind speed overlay with temp.
```{r}
ggplot(CTD_buoy_plot, aes(date,value)) +
  geom_point() +
  facet_grid(data_type ~., scales = "free") +
  theme_bw() +
  ggtitle("2016 Stillwater CTD and M1 buoy data (daily averages)") +
  scale_x_date(limits = as.Date(c('2016-06-01','2016-08-15'))) # set xlims

```
Save plot
```{r}
ggsave("StillwaterCTD_M1.pdf", width = 6, height = 8)
```
